Sterile Neutrinos, a Summary

### 2015-04-15 p3_505/presentation/pres.tex
typos

\begin{frame}
  \frametitle{Flavor Oscillations}
  \begin{itemize}
  \item Using $E\approx p$ (ultrarelativistic neutrinos), and $L = ct$:
    \begin{align}
      E_j - E_k &= \left(m_j^2 - m_k^2\right)/2E = \Delta m_{jk}^2/2E \\
      P &= \sum_{j=1}^3\sum_{k=1}^3 U^*_{\alpha j}U_{\alpha k}U_{\beta j}U_{\beta k}^* e^{-i\Delta m_{jk}^2L/2E}
    \end{align}
  \item And therefore a complete expression without hte complex exponential is:
    \begin{align}
        P(\nu_{\alpha}\rightarrow\nu_{\beta}) = \delta_{\alpha\beta} & - 4\sum_{j>k}\Re\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin^2\left(\frac{\Delta m_{jk}^2}{4E}L\right) \nonumber \\
  & + 2\sum_{j>k}\Im\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin\left(\frac{\Delta m_{jk}^2}{2E}L\right).
    \end{align}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{The LSND Experiment}
  \footnotesize{
    \begin{columns}
      \begin{column}{.5\linewidth}      
        \begin{itemize}
        \item Liquid Scintillator Neutrino Detector.
        \item Los Alamos National Laboratory from 1993 to 1998.
        \item Cherenkov detector using 1220 PMTs.
        \item Protons strike target, use pions decaying at test to product $\overline{\nu}_\mu$.
        \end{itemize}
      \end{column}
      \begin{column}{.5\linewidth}
        \begin{itemize}
        \item $117.9 \pm 22.4$ events, signal excess of $87.9\pm 22.4 \pm 6.0$.
        \item $(\sin^2 2\theta,\Delta m^2)$ fit extracts allowed oscillation regions.
        \item Best fits yields $\Delta m^2 \sim 1$ eV.
        \item Disagrees with solar and atmospheric oscillation experiments and also LEP experiments for 3 light flavor neutrinos.
        \end{itemize}
      \end{column}
    \end{columns}
  }
  \begin{center}
    \includegraphics[width=.35\linewidth]{../figures/LSND_plot.pdf}
  \end{center}
\end{frame}
\begin{frame}
  \frametitle{The MiniBooNE Experiment}
  \scriptsize{
    \begin{columns}
      \begin{column}{.5\linewidth}      
        \begin{itemize}
        \item Developed to answer the LSND anomaly.
        \item Collected data at Fermilab from 2002 to 2012.
        \item Cherenkov detector using PMTs (just like LSND).
        \item 8 GeV protons from the Booster strike beryllium target, pions decay producing neutrinos.
          \begin{center}
            \includegraphics[width=.9\linewidth]{../figures/miniboone_excess.pdf}
          \end{center}
        \end{itemize}
      \end{column}
      \begin{column}{.5\linewidth}
        \begin{itemize}
        \item Consistent with the LSND antineutrino oscillation result.
          \begin{center}
            \includegraphics[width=.6\linewidth]{../figures/miniboone_regions.pdf}
          \end{center}
        \item Detector systematics make it difficult to differentiate between $\pi^0\rightarrow\gamma\gamma$ and an electron in the detector.
        \end{itemize}
      \end{column}
    \end{columns}
  }
\end{frame}

@@

\begin{frame}
  \frametitle{(Selected) Future Experiments}
  \begin{itemize}
  \item MicroBooNE: Larger follow up experiment to MiniBooNE
   \begin{itemize}
     \item Time projection chamber with a 60 ton fiducial volume of liquid argon
     \item Will be able to distinguish between photons and electrons
     \item Expected to resolve the MiniBooNE excess anomaly
   \end{itemize}
  \item NOvA: Long baseline experiment using the NuMI beam
   \begin{itemize}
     \item 330 ton near and 14 kton far liquid scintillator detectors located near the surface
     \item Primary physics goals are oscillation measurements, will also perform sterile neutrino measurements
   \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
  \item A discovery of heavy, right-handed sterile neutrinos would have a large impact on our understanding of neutrino physics
  \item Some, very contested, evidence for sterile neutrinos has been found at LSND and MiniBooNE
  \item MINOS+, Super-K, Daya Bay, MicroBooNE, and NoVA will continue searching for sterile neutrinos
  \item Expect to see the anomalies be confirmed or refuted in the next year or two
  \end{itemize}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Flavor Oscillations}
  \begin{itemize}
  \item Using $E\approx p$ (ultrarelativistic neutrinos), and $L = ct$:
    \begin{align}
      E_j - E_k &= \left(m_j^2 - m_k^2\right)/2E = \Delta m_{jk}^2/2E \\
      P &= \sum_{j=1}^3\sum_{k=1}^3 U^*_{\alpha j}U_{\alpha k}U_{\beta j}U_{\beta k}^* e^{-i\Delta m_{jk}^2L/2E}
    \end{align}
  \item And therefore a complete expression without the complex exponential is:
    \begin{align}
        P(\nu_{\alpha}\rightarrow\nu_{\beta}) = \delta_{\alpha\beta} & - 4\sum_{j>k}\Re\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin^2\left(\frac{\Delta m_{jk}^2}{4E}L\right) \nonumber \\
  & + 2\sum_{j>k}\Im\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin\left(\frac{\Delta m_{jk}^2}{2E}L\right).
    \end{align}
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{The LSND Experiment}
  \footnotesize{
    \begin{columns}
      \begin{column}{.5\linewidth}      
        \begin{itemize}
        \item Liquid Scintillator Neutrino Detector.
        \item Los Alamos National Laboratory from 1993 to 1998.
        \item Cherenkov detector using 1220 PMTs.
        \item Protons strike target, use pions decaying at rest to product $\overline{\nu}_\mu$.
        \end{itemize}
      \end{column}
      \begin{column}{.5\linewidth}
        \begin{itemize}
        \item $117.9 \pm 22.4$ events, signal excess of $87.9\pm 22.4 \pm 6.0$.
        \item $(\sin^2 2\theta,\Delta m^2)$ fit extracts allowed oscillation regions.
        \item Best fits yields $\Delta m^2 \sim 1$ eV.
        \item Disagrees with solar and atmospheric oscillation experiments and also LEP experiments for 3 light flavor neutrinos.
        \end{itemize}
      \end{column}
    \end{columns}
  }
  \begin{center}
    \includegraphics[width=.35\linewidth]{../figures/LSND_plot.pdf}
  \end{center}
\end{frame}
\begin{frame}
  \frametitle{The MiniBooNE Experiment}
  \scriptsize{
    \begin{columns}
      \begin{column}{.5\linewidth}      
        \begin{itemize}
        \item Developed to answer the LSND anomaly.
        \item Collected data at Fermilab from 2002 to 2012.
        \item Cherenkov detector using PMTs (just like LSND).
        \item 8 GeV protons from the Booster strike beryllium target, pions decay producing neutrinos.
          \begin{center}
            \includegraphics[width=.9\linewidth]{../figures/miniboone_excess.pdf}
          \end{center}
        \end{itemize}
      \end{column}
      \begin{column}{.5\linewidth}
        \begin{itemize}
        \item Consistent with the LSND antineutrino oscillation result.
          \begin{center}
            \includegraphics[width=.6\linewidth]{../figures/miniboone_regions.pdf}
          \end{center}
        \item Detector systematics make it difficult to differentiate between $\pi^0\rightarrow\gamma\gamma$ and an electron in the detector.
        \end{itemize}
      \end{column}
    \end{columns}
  }
\end{frame}

@@

\begin{frame}
  \frametitle{(Selected) Future Experiments}
  \begin{itemize}
  \item MicroBooNE: Larger follow up experiment to MiniBooNE
   \begin{itemize}
     \item Time projection chamber with a 60 ton fiducial volume of liquid argon
     \item Will be able to distinguish between photons and electrons
     \item Expected to resolve the MiniBooNE excess anomaly
   \end{itemize}
  \item NO$\nu$A: Long baseline experiment using the NuMI beam
   \begin{itemize}
     \item 330 ton near and 14 kton far liquid scintillator detectors located near the surface
     \item Primary physics goals are oscillation measurements, will also perform sterile neutrino measurements
   \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Conclusions}
  \begin{itemize}
  \item A discovery of heavy, right-handed sterile neutrinos would have a large impact on our understanding of neutrino physics
  \item Some, very contested, evidence for sterile neutrinos has been found at LSND and MiniBooNE
  \item MINOS+, Super-K, Daya Bay, MicroBooNE, and NO$\nu$A will continue searching for sterile neutrinos
  \item Expect to see the anomalies be confirmed or refuted in the next year or two
  \end{itemize}
\end{frame}
